Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-01-27 paper/main.tex
.

\subsection{High level description}
To obtain a prediction for a given drug-protein pair, we first build DL2vec representations for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations. Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually. These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process. Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs. The overall architecture can be seen in Figure~\ref{fig:Fullmodel}.

@@

\begin{enumerate}
	\item Build split over drugs
	\item Build split over drug-target pairs
	\item Build split over proteins
\end{enumerate} 
In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results. This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}. The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias. Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive. As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes. \\

>>>>

\subsection{High level description}
Our aim is to develop and evaluate a computational model to predict
drug--target interactions. Specifically, given a protein, we are
interested in identifying the drugs that may target the protein, so as
to reflect the case where a drug is desired to target a
disease-associated protein.

To obtain a prediction for a given drug-protein pair, we first build DL2vec representations for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations. Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually. These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process. Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs. The overall architecture can be seen in Figure~\ref{fig:Fullmodel}.

@@

\todo[inline]{Relate these to the issues listed above:}
\begin{enumerate}
	\item Build split over drugs
	\item Build split over drug-target pairs
	\item Build split over proteins
\end{enumerate} 
In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results. This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}. The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias. Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive. As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes. \\
